Estimating the dynamics of SPARQL query results

### 2019-04-11 kgswc19.tex
update

\paragraph{Architecture:} In Figure~\ref{fig:schema}, we provide an overview of our proposed architecture for making predictions with respect to the OSC and TTL tasks. A SPARQL query $Q$ and a dynamic RDF graph $\mathcal{G}$ are given as input. The first component parses the query and generates static features which include statistics about the number of elements in the query (e.g., number of triple patterns, variables, etc.) as well as the presence of particular query operators (e.g., \texttt{UNION}, \texttt{FILTER}, recursive path expressions, etc.). The second component may take as input some features of the query, as well as information from the dynamic graph, and produce a set of further features; this component may compute, for example, statistics about the dynamicity of predicates used in the query, or statistics about changes in the query results over past versions. The final features are passed to a classifier in the case of the OSC problem, and to a regression model in the case of the TTL problem, to generate the final output prediction. We will now describe in more detail the query and data features considered in this work.

@@

In order to evaluate the effectiveness of our proposal for predicting changes in SPARQL query results, we require a dynamic RDF graph, with access to various historical versions; preferably this graph contains real-world, large-scale, diverse data, and with sufficient changes between versions to provide both positive and negative examples of queries whose results change. Furthermore, we require a set of SPARQL queries that can be answered against this dynamic graph; preferably these queries again should be diverse, representative of real-world user queries, of a variety of shapes and sizes, using a variety of query operators, and with a mix of both dynamic and static results over the dynamic graph. In particular, we choose Wikidata~\cite{VrandecicK14} for our experiments, which, we shall argue, meets the aforementioned requirements. We first give details of the dynamic data we collect from Wikidata; thereafter, we discuss the queries we use in our experiments.

@@

Although we have seen that the number of unique predicates did not change much over all versions, we are interested to see which predicates are indicative of dynamic statements. Table~\ref{tab:predicates} shows the ten most dynamic predicates according to the number statements involving that predicate (left) and the ratio of added (+) and deleted (−) statements divided by the total number of statements for that predicate across all snapshots; for the second, we only include predicates that appear in all snapshots and appear in ≥ 1, 000 statements overall. Note that the largest number of changes involves the predicate scheme:description, however, its dynamic value is low due to the number of statements with it in the versions.

@@

For OSC prediction, we selected four classifiers for testing: Decision Trees, Linear SVM, Naive Bayes and Nearest Neighbors. We split the data to use 80\% for training and 20\% for tests. To avoid overfitting, we use stratified cross-validation in the training data. We consider three types of features: static query features (\textsc{q}): predicate dynamics (\textsc{p}) considering information of $\Delta(\mathcal{G},p)$ and the cardinality of predicates and information on historical results (\texttt{r}) considering both, the number of past changes in the fixed interval and the cardinalities of those changes divided by the sizes of query results. We may also consider combinations of these features: \textsc{qp}, \textsc{qr}, \textsc{pr} and \textsc{qpr}.

>>>>

\paragraph{Architecture:} In Figure~\ref{fig:schema}, we provide an overview of a general architecture for making predictions with respect to the OSC and TTL tasks. A SPARQL query $Q$ and a dynamic RDF graph $\mathcal{G}$ are given as input. The first component parses the query and generates static features which include statistics about the number of elements in the query (e.g., number of triple patterns, variables, etc.) as well as the presence of particular query operators (e.g., \texttt{UNION}, \texttt{FILTER}, recursive path expressions, etc.). The second component may take as input some features of the query, as well as information from the dynamic graph, and produce a set of further features; this component may compute, for example, statistics about the dynamicity of predicates used in the query, or statistics about changes in the query results over past versions. The final features are passed to a classifier in the case of the OSC problem, and to a regression model in the case of the TTL problem, to generate the final output prediction. We will now describe in more detail the query and data features considered in this work.

@@

In order to evaluate the effectiveness for predicting changes in SPARQL query results, we require a dynamic RDF graph, with access to various historical versions; preferably this graph contains real-world, large-scale, diverse data, and with sufficient changes between versions to provide both positive and negative examples of queries whose results change. Furthermore, we require a set of SPARQL queries that can be answered against this dynamic graph; preferably these queries again should be diverse, representative of real-world user queries, of a variety of shapes and sizes, using a variety of query operators, and with a mix of both dynamic and static results over the dynamic graph. In particular, we choose Wikidata~\cite{VrandecicK14} for our experiments, which, we shall argue, meets the aforementioned requirements. We first give details of the dynamic data we collect from Wikidata; thereafter, we discuss the queries we use in our experiments.

@@

Although we have seen that the number of unique predicates did not change much over all versions, we are interested to see which predicates are indicative of dynamic statements. Table~\ref{tab:predicates} shows the ten most dynamic predicates according to the number of added and deleted statements involving that predicate (Total) and the ratio of added and deleted statements divided by the total number of statements for that predicate across all snapshots (Dyn); for the second, we only include predicates that appear in all snapshots and appear in ≥ 1, 000 statements overall. Note that the largest number of changes involves the predicate scheme:description, however, its dynamic value is low due to the number of statements with it in the versions.

@@

For OSC prediction, we selected four classifiers for testing: Decision Trees, Linear SVM, Naive Bayes and Nearest Neighbors. We split the data to use 80\% for training and 20\% for tests. To avoid overfitting, we use stratified cross-validation in the training data. We consider three types of features: static query features (\textsc{q}): predicate dynamics (\textsc{p}) considering information of $\Delta(\mathcal{G},p)$ and the cardinality of predicates and information on historical results (\textsc{r}) considering the number of past changes in the fixed interval. We may also consider combinations of these features: \textsc{qp}, \textsc{qr}, \textsc{pr} and \textsc{qpr}.
